Typed Lua

### 2013-06-24 doc/slides/gsoc_proposal/gsoc_proposal.tex
Added an example of gradual typing to the presentation about the GSoC proposal

\title{Typed Lua}
\subtitle{Google Summer of Code Proposal}
\author{André Murbach Maidl}
\institute[PUC-Rio]{LabLua\\PUC-Rio}
\date{}

@@

\subsection{The levels of Gradual Typing}
\begin{frame}
\frametitle{The levels of Gradual Typing}
\begin{itemize}
\item The programmer can gradually migrate to static typing
\item Specially in module boundaries
\begin{enumerate}
\item The programmer write modules that use only dynamic typing.
\item The programmer creates an interface to type a module that
is written using only dynamic typing. This module is not
type checked, but their code that use this module is.
\item The programmer write modules that use only static typing.
\end{enumerate}
\end{itemize}
\end{frame}

\subsection{Gradual Typing and Dynamically Typed Languages}
\begin{frame}
\frametitle{Gradual Typing and Dynamically Typed Languages}
\begin{itemize}
\item Dart and TypeScript
\begin{itemize}
\item New languages derived from JavaScript
\item Both target JavaScript
\item Both are designed with a gradual type system
\end{itemize}
\item Typed Racket and core.typed
\begin{itemize}
\item Gradual typing extensions to Racket and Clojure
\end{itemize}
\item Perl 6
\begin{itemize}
\item Not compatible with Perl 5
\end{itemize}
\item Python
\begin{itemize}
\item gradual (package)
\item Runtime type checking semantics of gradual typing
\item Works only on Python 3
\end{itemize}
\end{itemize}
\end{frame}

@@

\subsection{Related Work}
\begin{frame}
\frametitle{Related Work}
\begin{itemize}
\item Some Lua projects implement parts of what is being proposed
\item None implement type checking along with OO programming
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{LuaInspect}
\begin{itemize}
\item A tool that performs code analysis
\item For instance, flagging undeclared identifiers
\item It does not try to analyze OO code
\item It does not perform type checking
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{MetaLua and MoonScript}
\begin{itemize}
\item Both support some sort of OO programming
\item Both do not perform type checking
\end{itemize}
\end{frame}

@@

\subsection{Some Statistics}
\begin{frame}
\frametitle{Some Statistics}
\begin{itemize}
\item Data from LuaRocks repository as of June 10th
\item LuaRocks repository contains 2950 .lua files
\item \textbf{luac} (version 5.2.2) compiles 2864 of the 2950 files
\item We have 2069 files when we ignore files that are generated by
MoonScript and documentation and test subdirectories
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Function declaration}
\begin{itemize}
\item 18864 functions are being declared
\begin{itemize}
\item 4127 (21.9\%) are actually methods
\item 1586 (8.4\%) use \textbf{type} to inspect the tags of the arguments
\item 771 (4.1\%) return \textbf{nil} plus something else
\item 119 (0.6\%) return \textbf{false} plus something else
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Metatables}
\begin{itemize}
\item 18864 functions are being declared
\begin{itemize}
\item 549 (2.9\%) use \textbf{setmetatable}
\item 175 (0.9\%) use \textbf{getmetatable}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Table constructor}
\begin{itemize}
\item The table constructor is being used 19049 times
\begin{itemize}
\item 4571 (24.00\%) times is the empty construct
\item 14475 (75.99\%) times is a constructor with only static keys
\item 3 (0.01\%) times is a constructor with only dynamic keys
\item There is no constructor with static and dynamic keys
\end{itemize}
\end{itemize}
\end{frame}

>>>>

\title{Typed Lua}
\subtitle{Google Summer of Code Proposal}
\author{André Murbach Maidl}
\institute[PUC-Rio]{LabLua\\PUC-Rio}
\date{June 27th 2013}

@@

\begin{frame}[fragile]
\frametitle{Gradual Typing Example}
\begin{verbatim}
local function div (x, y)
  if y == 0 then
    return nil,"division by zero"
  end
  return x / y
end

local function div (x:Number, y:Number) -> Any
  if y == 0 then
    return nil,"division by zero"
  end
  return x / y
end
\end{verbatim}
\end{frame}

\subsection{The levels of Gradual Typing}
\begin{frame}
\frametitle{The levels of Gradual Typing}
\begin{itemize}
\item The programmer can gradually migrate to static typing
\item Specially in module boundaries
\begin{enumerate}
\item The programmer write modules that use only dynamic typing.
\item The programmer creates an interface to type a module that
is written using only dynamic typing. This module is not
type checked, but the code that uses this module is.
\item The programmer write modules that use only static typing.
\end{enumerate}
\end{itemize}
\end{frame}

\subsection{Gradual Typing and Dynamically Typed Languages}
\begin{frame}
\frametitle{Gradual Typing and Dynamically Typed Languages}
\begin{itemize}
\item Dart and TypeScript
\begin{itemize}
\item New languages derived from JavaScript
\item Both target JavaScript
\item Both are designed with a gradual type system
\end{itemize}
\item Typed Racket and core.typed
\begin{itemize}
\item Gradual typing extensions to Racket and Clojure
\end{itemize}
\item Perl 6
\begin{itemize}
\item Not compatible with Perl 5
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Gradual Typing and Dynamically Typed Languages}
\begin{itemize}
\item Python
\begin{itemize}
\item py-gradual
\begin{itemize}
\item A package that provides runtime semantics of gradual typing
\item Works only on Python 3
\end{itemize}
\item reticulated
\begin{itemize}
\item A tool that implements gradual typing
\item Works on Python 2.7, 3.2, and 3.3
\end{itemize}
\end{itemize}
\item Jython
\begin{itemize}
\item jython-gradual
\begin{itemize}
\item A fork that includes gradual typing
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\subsection{Related Work}
\begin{frame}
\frametitle{Related Work}
\begin{itemize}
\item Some Lua projects implement parts of what is being proposed
\item None implement compile-time type checking along with OO programming
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{MetaLua and MoonScript}
\begin{itemize}
\item Both support some sort of OO programming
\item Both do not perform compile-time type checking
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{LuaInspect}
\begin{itemize}
\item A tool that uses MetaLua to perform code analysis
\begin{itemize}
\item Flags unknown global variables and table fields
\item Checks number of function arguments against signatures
\item Infers function return values
\end{itemize}
\item It does not try to analyze OO code
\item It does not perform compile-time type checking
\end{itemize}
\end{frame}

@@

\section{Statistics}
\begin{frame}[fragile]
\frametitle{Statistics}
\begin{itemize}
\item Data from LuaRocks repository as of June 22nd
\item LuaRocks repository contains 3000 \verb'.lua' files
\item \textbf{luac} (version 5.2.2) compiles 2913 of the 3000 files
\item The statistics include 1966 of the 2913 files
\begin{itemize}
\item I did not include files that are generated by MoonScript
\item I did not include documentation subdirectories
\item I did not include test subdirectories
\item I did not include test files in general
\end{itemize}
\end{itemize}
\end{frame}

\subsection{Function declaration}
\begin{frame}
\frametitle{Function declaration}
\begin{itemize}
\item 17390 functions are being declared in the 1966 files
\begin{itemize}
\item 3961 (22.8\%) are actually methods
\item 1512 (8.7\%) use \textbf{type} to inspect the tags of the arguments
\item 773 (4.5\%) return \textbf{nil} plus something else
\item 118 (0.7\%) return \textbf{false} plus something else
\end{itemize}
\end{itemize}
\end{frame}

\subsection{Metatables}
\begin{frame}
\frametitle{Metatables}
\begin{itemize}
\item 294 (15.0\%) of the 1966 files use \textbf{setmetatable}
\item 99 (5.1\%) of the 196 files use \textbf{getmetatable}
\end{itemize}
\end{frame}

\subsection{Table constructor}
\begin{frame}
\frametitle{Table constructor}
\begin{itemize}
\item The table constructor is being used 17369 times in the 1966 files
\begin{itemize}
\item 4391 (25.28\%) times is the empty construct
\item 12975 (74.70\%) times is a constructor with only static keys
\item 3 (0.02\%) times is a constructor with only dynamic keys
\item There is no constructor with static and dynamic keys
\end{itemize}
\end{itemize}
\end{frame}

\subsection{Module}
\begin{frame}
\frametitle{Module}
\begin{itemize}
\item 655 (33.3\%) of the 1966 files use \textbf{return} at the end of the file
\item 602 (30.7\%) of the 1966 files use the function \textbf{module}
\item The rest of the files are plain scripts
\end{itemize}
\end{frame}
